The Ontological Modal Collapse \\ as a Collapse of the Square of Opposition

### 2014-09-08 Papers/Universalis/ModalCollapse.tex
Write conclusion

All results announced in this note have been obtained experimentally\footnote{The source code of the experiments are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.} using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}.

Various slightly different versions of
axioms and definitions have been considered by G\"{o}del and by
several philosophers who commented on his proof
(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).

In theoretical philosophy, formal logical confrontations with such
ontological arguments had been so far (mainly) limited to paper
and pen.  Up to now, the use of computers was prevented, because the
logics of the available theorem proving systems were not expressive
enough to formalize the abstract concepts adequately. G{\"o}del's proof
uses, for example, a complex higher-order modal logic (\HOML) to handle
concepts such as \emph{possibility} and \emph{necessity} and to support
quantification over individuals and properties.

controversies, care with parameters

ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}

The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist . This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.

>>>>

All results announced in this note have been obtained experimentally using interactive and automated theorem provers and model finders \cite{ToDo:leo,sattalax,isabelle,coq,nitpick}. The source codes of the experiments, as well as the resulting proofs and counter-models, are available in \url{https://github.com/FormalTheology/GoedelGod/tree/master/Formalizations/Isabelle}.

The technique enabling these experiments is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist. This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description is available in \cite{todo:ECAI}.

The modal collapse is an interesting example of philosophical controversy and dispute, to which Leibniz's idea of a \emph{calculus ratiocinator} in the form of contemporary automated theorem provers can be applied. Current technology is increasingly ready to be embraced by those willing to practice computer-assisted theoretical philosophy \cite{oppenheimera11,rushby13}.

% Various slightly different versions of
% axioms and definitions have been considered by G\"{o}del and by
% several philosophers who commented on his proof
% (cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).

% In theoretical philosophy, formal logical confrontations with such
% ontological arguments had been so far (mainly) limited to paper
% and pen.  Up to now, the use of computers was prevented, because the
% logics of the available theorem proving systems were not expressive
% enough to formalize the abstract concepts adequately. G{\"o}del's proof
% uses, for example, a complex higher-order modal logic (\HOML) to handle
% concepts such as \emph{possibility} and \emph{necessity} and to support
% quantification over individuals and properties.

% controversies, care with parameters

% ToDo: Leibniz calculemus, Rushby, Zalta
